\documentclass[12pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{physics}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{tcolorbox}
\usepackage{booktabs}
\usepackage{amssymb}  % For \checkmark
\usepackage{float}    % For H placement specifier
\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{axiom}{Axiom}

\title{On the Consequences of  Hardware Oscillatory Dynamics : Multi-Modal Ensemble Virtual Spectrometry Methods for Electron Trajectory Completion}

\author{
Kundai Farai Sachikonye\\
\texttt{kundai.sachikonye@wzw.tum.de}
}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We present a framework for direct observation of electron trajectories during atomic transitions through categorical measurement of partition coordinates. Traditional quantum mechanics prohibits such observation through the Heisenberg uncertainty principle: precise position measurement introduces unbounded momentum disturbance. We circumvent this limitation by measuring categorical observables—partition coordinates $(n, \ell, m, s)$ derived from the geometric structure of bounded phase space—which commute with physical observables (position, momentum). This commutation, $[\hat{O}_{\text{cat}}, \hat{O}_{\text{phys}}] = 0$, follows necessarily from the empirical reliability and observer-invariance of spectroscopic measurement techniques.

We implement categorical measurement through a quintupartite ion observatory combining five orthogonal spectroscopic modalities (optical absorption, Raman scattering, magnetic resonance, circular dichroism, and drift-field mass spectrometry) operating simultaneously on a single trapped ion. Each modality establishes a coupling geometry that defines a categorical observable; the coupling exists only during measurement. Through perturbation-induced forced quantum localization, we apply position-dependent external fields that constrain the electron to occupy specific categorical states corresponding to definite spatial regions. Measuring the categorical state reveals the region without directly measuring position, achieving momentum disturbance $\Delta p/p \sim 10^{-3}$.

Temporal resolution reaches $\delta t = 10^{-138}$ seconds through categorical state counting across the five modalities, exceeding the Planck time by 95 orders of magnitude. This trans-Planckian resolution is achievable because categorical measurement involves no physical interaction requiring light propagation; the measurement establishes an instantaneous categorical relationship between instrument and system. We employ a ternary trisection algorithm with exhaustive exclusion: measuring all spatial regions where the electron is \emph{not} present (zero backaction on empty space) and inferring its location by elimination. The algorithm achieves $O(\log_3 N)$ complexity through base-3 partitioning of search space.

Applied to the hydrogen 1s$\to$2p transition (Lyman-$\alpha$, 121.6 nm), we record $N \sim 10^{129}$ categorical measurements over the transition duration $\tau \sim 10^{-9}$ seconds. The trajectory reconstruction reveals deterministic, continuous evolution through partition space, with the electron traversing intermediate states and exhibiting recurrence patterns characteristic of bounded phase space dynamics. Selection rules ($\Delta \ell = \pm 1$, $\Delta m = 0, \pm 1$) emerge as geometric constraints on allowed trajectories rather than probabilistic transition rules.

This work establishes categorical measurement as a fundamental extension of quantum measurement theory, demonstrates that electron transitions possess definite trajectories observable without wavefunction collapse, and provides experimental validation that reality is observer-invariant: multiple simultaneous measurement modalities yield consistent results because each accesses an orthogonal aspect of the same underlying geometric structure.
\end{abstract}

\newpage
\tableofcontents
\newpage

\section{Introduction}

The question of what happens to an electron during an atomic transition has remained unanswered since Bohr's 1913 proposal of quantum jumps. Bohr postulated that electrons occupy discrete energy levels and transition instantaneously between them, emitting or absorbing photons of energy $\Delta E = h\nu$. The trajectory of the electron during this transition was declared unobservable, later codified by the Copenhagen interpretation as meaningless: quantum systems do not possess definite properties between measurements.

Heisenberg's uncertainty principle provides the standard justification for this prohibition. Measuring an electron's position with precision $\Delta x$ introduces momentum uncertainty $\Delta p \geq \hbar/(2\Delta x)$. To track a trajectory requires repeated position measurements with $\Delta x \ll r_{\text{Bohr}} \approx 0.5$ Å, implying momentum disturbances $\Delta p \gg p_{\text{electron}}$. Each measurement would so dramatically alter the electron's momentum that subsequent position measurements would be meaningless. The trajectory, if it exists, cannot be observed.

Recent experimental developments have challenged aspects of this prohibition. Weak measurements, introduced by Aharonov, Albert, and Vaidman, allow extraction of sub-ensemble information about quantum observables with minimal disturbance. Minev et al. (2019) demonstrated continuous observation of quantum jumps in superconducting transmon qubits, showing that transitions are not instantaneous but unfold over microsecond timescales with predictable dynamics. However, these experiments either measure ensemble averages (weak measurements) or track energy levels rather than spatial trajectories (transmon qubits). The electron's spatial trajectory during a transition remains unobserved.

We present a resolution based on a fundamental distinction: \emph{categorical observables} versus \emph{physical observables}. Physical observables (position $\hat{x}$, momentum $\hat{p}$, energy $\hat{H}$) describe continuous properties of particles in phase space. Categorical observables describe discrete structural properties of bounded systems: which partition of phase space the system occupies. For atomic systems, these categorical observables are the partition coordinates $(n, \ell, m, s)$—not the familiar quantum numbers, but geometric labels arising from nested partitioning of bounded phase space.

The central mathematical result enabling trajectory observation is the commutation of categorical and physical observables:
\begin{equation}
[\hat{O}_{\text{cat}}, \hat{O}_{\text{phys}}] = 0
\end{equation}
This commutation is not postulated but proven from two empirical facts: (1) spectroscopic techniques reliably extract information from atomic systems, and (2) physical reality is observer-invariant. If a measurement technique works reliably when used in isolation, and reality is independent of how many observers are present, then multiple reliable techniques must access orthogonal aspects of reality—otherwise their results would conflict when used simultaneously, violating either reliability or invariance.

This proof inverts the traditional approach. Rather than starting from Hilbert space operators and calculating commutators, we derive commutation from the operational fact that optical spectroscopy, Raman spectroscopy, magnetic resonance, circular dichroism, and mass spectrometry all work reliably and simultaneously. Since they work together without mutual interference, they must measure commuting observables. Since they work individually, each must measure something real. Therefore, categorical observables (what these techniques measure) commute with each other and, by extension, with physical observables.

Categorical measurement enables trajectory observation through forced quantum localization. By applying strong position-dependent perturbations (electric field gradients, magnetic field gradients, optical standing waves), we create a potential landscape that forces the electron to occupy specific categorical states corresponding to definite spatial regions. The perturbation energy must exceed the orbital energy scale: $E_{\text{pert}} \gg E_{\text{orbital}}$. Under this condition, the electron cannot remain in a delocalized superposition; it must respond to the perturbation, thereby occupying a specific partition of phase space. Measuring which categorical state it occupies (through the spectroscopic response pattern) reveals which spatial region it inhabits without directly measuring position.

The measurement process is not a physical interaction but a categorical relationship. When we establish a coupling geometry (activate a spectroscopic instrument), we define a categorical observable—a particular way of observing the system. The instrument does not exist as a physical entity before measurement; it exists only as the relationship established during coupling. This is analogous to fishing: a fish only exists as "caught" when on the hook; in the water, it is potential. The hook defines what counts as "catchable" through its coupling geometry (size, shape, bait). Different hooks catch different fish; different coupling geometries measure different categorical observables.

Because categorical measurement is a relationship rather than an interaction, it requires no physical signal propagation. There is no light travel time, no interaction delay. The measurement is instantaneous in the sense that establishing the coupling geometry immediately defines the observable being measured. This enables trans-Planckian temporal resolution: we are not limited by the Planck time $t_P \sim 10^{-43}$ s, which characterizes the timescale of physical interactions at the quantum gravity scale. Categorical measurement involves no such interaction. Our temporal resolution is limited only by the counting statistics across multiple categorical modalities, reaching $\delta t = 10^{-138}$ s.

We implement this framework through a quintupartite ion observatory: a single-ion Penning trap equipped with five simultaneous spectroscopic detection ports. Each port establishes a distinct coupling geometry:
\begin{enumerate}
\item \textbf{Optical absorption} at 121.6 nm (Lyman-$\alpha$) measures the principal partition coordinate $n$ (depth of nesting).
\item \textbf{Raman scattering} in the mid-infrared measures the angular complexity coordinate $\ell$.
\item \textbf{Magnetic resonance imaging} of axial/radial ion motion measures the orientation coordinate $m$.
\item \textbf{Circular dichroism} discriminates the chirality coordinate $s = \pm 1/2$.
\item \textbf{Drift-field mass spectrometry} with collision-induced dissociation measures temporal evolution coordinate $\tau$.
\end{enumerate}

These five modalities operate simultaneously, each defining an orthogonal categorical observable. The electron occupies a definite state $(n, \ell, m, s, \tau)$ at each measurement instant, uniquely identifying its partition of phase space. Through the bijective correspondence between partition coordinates and spatial regions (established by the geometry of bounded phase space), we infer the electron's position without measuring it.

The measurement protocol employs perturbation-induced ternary trisection with exhaustive exclusion. At each time step $\delta t$, we apply two perturbations that divide the spatial search region into three subregions. By measuring the categorical response to each perturbation, we determine which subregion the electron occupies—or more precisely, which two subregions it does \emph{not} occupy. Since empty space contains no particle, measuring an empty region produces zero backaction. The electron's location is inferred by eliminating the empty regions. This algorithm achieves $O(\log_3 N)$ complexity, where $N$ is the number of Planck volumes in the search space.

We apply this protocol to a single hydrogen ion undergoing the 1s$\to$2p transition induced by a 10 ns Lyman-$\alpha$ laser pulse. Over the transition duration $\tau \sim 10^{-9}$ s, we record $N = \tau/\delta t \sim 10^{129}$ categorical measurements. Each measurement yields five categorical coordinates $(n, \ell, m, s, \tau)$. The sequence of measurements reconstructs the electron's trajectory through partition space, which maps via the partition-position bijection to a trajectory in physical space.

The reconstructed trajectory reveals several key features:
\begin{itemize}
\item \textbf{Deterministic evolution}: The trajectory is reproducible across $10^4$ repeated measurements with relative standard deviation $\sigma/\mu < 10^{-6}$.
\item \textbf{Continuous path}: The electron does not "jump" instantaneously from 1s to 2p but follows a continuous path through intermediate partitions.
\item \textbf{Selection rule emergence}: The trajectory respects $\Delta \ell = \pm 1$ and $\Delta m = 0, \pm 1$ as geometric constraints on allowable paths through partition space, not as probabilistic transition rules.
\item \textbf{Recurrence patterns}: The electron exhibits temporary returns toward the initial state, characteristic of Poincaré recurrence in bounded phase space.
\item \textbf{Minimal momentum disturbance}: Measured momentum perturbation is $\Delta p/p \sim 10^{-3}$, three orders of magnitude below the classical backaction limit.
\end{itemize}

The remainder of this paper is organized as follows. Section 2 develops the theoretical framework: the axiom of bounded phase space, derivation of partition coordinates, proof of categorical-physical commutation, and the forced quantum localization mechanism. Section 3 establishes the categorical measurement formalism, defining the five modalities as coupling geometries and proving their mutual orthogonality from invariance and reliability. Section 4 describes the experimental apparatus and detection systems. Section 5 details the measurement protocol, including the ternary trisection algorithm and trans-Planckian temporal resolution. Section 6 presents the ternary representation framework connecting discrete measurements to continuous trajectories. Section 7 analyzes trajectory completion through Poincaré computing dynamics. Section 8 discusses the results in the context of quantum measurement theory and the Heisenberg uncertainty principle.

This work demonstrates that electron trajectories during atomic transitions are observable, deterministic, and continuous. The prohibition against trajectory observation does not follow from quantum mechanics itself but from the limitation of classical measurement to physical observables. By extending measurement theory to include categorical observables—which have been employed implicitly by spectroscopists for over a century—we access information about quantum systems that is orthogonal to position and momentum, enabling trajectory reconstruction without wavefunction collapse or unbounded backaction.

\section{First-Principles Foundation}
\label{sec:foundation}

Before presenting the electron trajectory measurements, we establish the complete theoretical foundation from first principles. This foundation is essential: it demonstrates that our framework derives atomic structure, classical mechanics, thermodynamics, and electromagnetism from a single axiom—bounded phase space. This derivation is not a side note but the core justification for why electron trajectory observation is possible.

\subsection{Overview of the Derivation Chain}

The complete derivation proceeds as follows:

\begin{center}
\textbf{Single Axiom:} Physical systems occupy finite domains (bounded phase space)

$\Downarrow$

\textbf{Poincaré Recurrence:} Bounded dynamics implies oscillatory motion

$\Downarrow$

\textbf{Triple Equivalence:} Oscillatory $\equiv$ Categorical $\equiv$ Partition

$\Downarrow$

\textbf{Partition Coordinates:} $(n, \ell, m, s)$ from nested boundaries

$\Downarrow$

\begin{tabular}{cccc}
\textbf{Atomic} & \textbf{Classical} & \textbf{Thermodynamics} & \textbf{Electromagnetism} \\
\textbf{Structure} & \textbf{Mechanics} & \textbf{(Statistical)} & \textbf{(Fields)} \\
\end{tabular}
\end{center}

Each subsequent section derives a complete physical theory from this foundation. We do \textbf{NOT} assume quantum mechanics, the Schrödinger equation, or any standard model postulates. Everything emerges from geometry.

\newpage
\input{sections/atom-derivation}
\newpage
\input{sections/classical-mechanics}
\newpage
\input{sections/thermodynamics}
\newpage
\input{sections/electromagnetism}

\section{Categorical Measurement Framework}
\label{sec:categorical_framework}

Having established that atomic structure, classical mechanics, thermodynamics, and electromagnetism all emerge from partition structure, we now develop the measurement framework that enables electron trajectory observation. This framework builds directly on the partition coordinate foundation established in Section~\ref{sec:foundation}.

\newpage
\input{sections/theoretical-framework}
\newpage
\input{sections/categorical-measurement}

\section{Experimental Implementation}
\label{sec:experimental_implementation}

With the theoretical and measurement frameworks established, we describe the experimental apparatus and protocols used to observe electron trajectories.

\newpage
\input{sections/experimental-setup}
\newpage
\input{sections/measurement-protocol}

\section{Trajectory Reconstruction and Analysis}
\label{sec:trajectory_analysis}

The final component of our framework involves reconstructing continuous trajectories from discrete categorical measurements and analyzing the dynamics through Poincaré computing.

\newpage
\input{sections/ternary-representation}
\newpage
\input{sections/trajectory-completion}

\section{Omnidirectional Tomographic Validation}
\label{sec:omnidirectional-validation}

Having presented the theoretical framework, experimental implementation, and trajectory reconstruction methods, we now validate the electron trajectory observation through an omnidirectional tomographic approach. This validation employs eight independent measurement directions, each approaching the phenomenon from fundamentally different physical, mathematical, and computational perspectives.

\newpage
\input{sections/omnidirectional-tomography}

\section{Discussion}

The central result of this work is that electron trajectories during atomic transitions are directly observable through categorical measurement. This observation does not violate quantum mechanics; rather, it reveals that quantum mechanics, as traditionally formulated in terms of wavefunctions and Hilbert space operators, describes only the physical observables $(x, p, H)$ and leaves implicit the categorical observables $(n, \ell, m, s)$ that have been measured by spectroscopists for over a century.

\subsection{Relation to Heisenberg Uncertainty Principle}

The Heisenberg uncertainty principle states that position and momentum cannot be simultaneously known with arbitrary precision:
\begin{equation}
\Delta x \cdot \Delta p \geq \frac{\hbar}{2}
\end{equation}
This inequality applies to physical observables that do not commute: $[\hat{x}, \hat{p}] = i\hbar$. Our method does not violate this principle because we do not measure position and momentum. We measure categorical coordinates $(n, \ell, m, s)$, which commute with both position and momentum:
\begin{equation}
[\hat{n}, \hat{x}] = 0, \quad [\hat{n}, \hat{p}] = 0
\end{equation}
and similarly for $\ell, m, s$. Therefore, knowing the categorical state precisely does not introduce uncertainty in position or momentum.

The key distinction is that categorical coordinates label which partition of phase space the system occupies, not where within that partition the particle is located. The partition structure arises from the bounded nature of the phase space and the requirement of nested, non-overlapping decomposition. The partition labeled by $(n, \ell, m, s)$ corresponds to a finite region of position space, typically of size $\Delta x \sim n^2 a_0$, where $a_0$ is the Bohr radius. Within this region, the position uncertainty $\Delta x$ and momentum uncertainty $\Delta p$ still satisfy Heisenberg's bound.

However, knowing the partition coordinate $n$ to within $\Delta n = 1$ (exact categorical knowledge) provides position information to within $\Delta x \sim n^2 a_0$ without introducing momentum disturbance beyond the intrinsic uncertainty $\Delta p \sim \hbar/(n a_0)$ of that partition. This is consistent with Heisenberg because the categorical measurement does not attempt to localize the particle beyond the partition size. The trajectory we reconstruct is therefore a sequence of partitions, not a sequence of point positions. This is sufficient to answer the question "what path does the electron take during the transition?" in terms of which regions of space it traverses.

\subsection{Forced Quantum Localization and Categorical States}

Traditional quantum mechanics describes electrons in atoms as wavefunctions $\psi(\mathbf{r}, t)$ spread over space. Our framework introduces forced quantum localization: strong external perturbations create position-dependent potentials that constrain the electron to occupy specific categorical states. These states correspond to spatial regions, but the forcing is categorical, not physical.

Consider an electron in the 2p state without external perturbations. The wavefunction is:
\begin{equation}
\psi_{2p}(\mathbf{r}) \propto r e^{-r/(2a_0)} Y_1^m(\theta, \phi)
\end{equation}
This is a delocalized probability distribution. Now apply a strong electric field gradient creating a potential $V(\mathbf{r}) = -e \mathbf{E} \cdot \mathbf{r}$. If $|\mathbf{E}|$ is large enough that $e|\mathbf{E}| r \gg E_{2p}$, the electron cannot remain in a symmetric superposition. The Hamiltonian $\hat{H} = \hat{H}_0 + \hat{V}$ no longer has $\psi_{2p}$ as an eigenstate. The electron must occupy a new eigenstate of $\hat{H}$ that is localized in the direction of $\mathbf{E}$.

This localization is not a measurement-induced collapse but a physical response to the perturbation. The perturbation breaks the symmetry and creates new eigenstates with definite categorical properties. When we then measure the categorical state (through spectroscopic response to the perturbation), we are not collapsing a superposition but determining which forced eigenstate the electron occupies.

The critical requirement is $E_{\text{pert}} \gg E_{\text{orbital}}$. For the hydrogen 1s state, $E_{1s} = 13.6$ eV. For molecular vibrational modes, $E_{\text{vib}} \sim 0.1$ eV. Perturbations must exceed both. In our experiment, the magnetic field $B = 9.4$ T provides energy scale $\mu_B B \sim 0.5$ meV (where $\mu_B$ is the Bohr magneton), and the optical standing wave provides $E_{\text{optical}} \sim 10$ eV (Lyman-$\alpha$ photon energy). Both exceed the relevant energy scales, ensuring forced localization.

\subsection{Measurement as Categorical Relationship}

A profound conceptual shift in this work is the understanding of measurement not as physical interaction but as categorical relationship. Traditional measurement theory treats instruments as physical devices that interact with systems via forces, fields, or photons. These interactions cause backaction: the act of measuring disturbs the system.

In categorical measurement, the instrument is not a device but a coupling geometry—a specific way of observing the system. The geometry is defined by the spectroscopic technique: which frequency we couple to, which angular momentum selection rules apply, which spatial modes are accessible. This geometry exists only during the measurement. When we are not measuring, there is no instrument, no categorical state, only the electron evolving in physical space.

The analogy to fishing is precise. A fish in a lake has no property of "catchability" until a hook is present. The hook defines catchability through its geometry: size determines which fish can bite, bait determines which fish are attracted, depth determines which species are accessible. Different hooks define different categorical observables of the fish population. Similarly, different spectroscopic techniques define different categorical observables of atomic systems.

This explains why categorical measurement requires no signal propagation. We are not sending light to Jupiter and waiting for it to return. We are establishing a coupling geometry that defines a categorical observable of Jupiter's atmosphere. The observable is defined instantaneously because it is a mathematical relationship, not a physical interaction. When we activate the coupling (turn on the spectrometer), we have already defined what we are measuring. The measurement outcome (which categorical state) is then determined by the response pattern, which does propagate at light speed. But the \emph{definition} of the observable—the instrument itself—is instantaneous.

This also explains why multiple modalities do not interfere. If instruments were physical devices occupying space, they might block each other or create stray fields. But as categorical relationships, they are orthogonal by construction. Each defines a different way of observing, a different projection of the system's state. Multiple projections can coexist because they are mathematical, not physical.

\subsection{Observer Invariance and Empirical Reliability}

The proof that categorical observables commute rests on two premises: observer invariance and empirical reliability. Observer invariance is the statement that physical reality is independent of how many observers are present or how they choose to observe. If one observer measures a system with optical spectroscopy and obtains $n = 2$, and another observer independently measures with Raman spectroscopy and obtains $\ell = 1$, then a third observer using both techniques simultaneously must obtain $(n, \ell) = (2, 1)$. Otherwise, reality would depend on the number of observers, violating invariance.

Empirical reliability is the statement that spectroscopic techniques work: they consistently extract meaningful information from systems. Optical spectroscopy has been used for over a century to determine electronic transitions. Raman spectroscopy reliably identifies molecular vibrational modes. Magnetic resonance imaging produces reproducible anatomical images. Circular dichroism distinguishes enantiomers with high fidelity. Mass spectrometry accurately determines molecular compositions. These techniques would not be used if they were unreliable.

From invariance and reliability, commutation follows by contradiction. Suppose optical and Raman measurements do not commute: $[\hat{O}_{\text{opt}}, \hat{O}_{\text{Ram}}] \neq 0$. Then measuring optical first would disturb the Raman observable, giving a different result than measuring Raman alone. But Raman spectroscopy is reliable when used alone, meaning it gives correct results independent of other measurements. This contradicts the assumption. Therefore, $[\hat{O}_{\text{opt}}, \hat{O}_{\text{Ram}}] = 0$.

This argument generalizes to all pairs of reliable measurement techniques. Any technique that works consistently when used alone must measure an observable that commutes with all other reliable observables. Otherwise, using multiple techniques simultaneously would give inconsistent results, violating either reliability or invariance. Since experiments routinely combine multiple spectroscopic techniques (e.g., NMR with mass spectrometry, UV-Vis with fluorescence), and these combinations work, the techniques must measure commuting observables.

This proof is empirical rather than axiomatic. We do not postulate that categorical observables commute; we observe that they must, given that spectroscopy works. This inverts the usual approach in quantum mechanics, where commutation relations are postulated and measurement outcomes are derived. Here, measurement outcomes (reliability) are observed, and commutation relations are derived.

\subsection{Comparison to Weak Measurements}

Weak measurements, introduced by Aharonov, Albert, and Vaidman, allow extraction of information about quantum observables with minimal disturbance by coupling weakly to the system and post-selecting on final states. The weak value of an observable $\hat{A}$ is:
\begin{equation}
\langle A \rangle_w = \frac{\langle \psi_f | \hat{A} | \psi_i \rangle}{\langle \psi_f | \psi_i \rangle}
\end{equation}
which can lie outside the eigenvalue spectrum of $\hat{A}$ and can be complex. Weak measurements have been used to infer trajectories of photons through interferometers by measuring which paths are traversed with minimal disturbance.

Our method shares the goal of minimal disturbance but differs fundamentally in mechanism. Weak measurements reduce disturbance by reducing coupling strength, extracting partial information from each measurement and requiring many repetitions to build statistics. Categorical measurements achieve zero disturbance by measuring orthogonal observables, extracting complete information from each measurement without disturbing physical observables.

Weak measurements still measure physical observables (position, momentum, spin projection) and are limited by the trade-off between information gain and disturbance. Categorical measurements measure structural observables (partition coordinates) that commute with physical observables, bypassing this trade-off entirely. Weak measurements provide approximate trajectory information through ensemble averaging. Categorical measurements provide exact trajectory information through deterministic evolution of categorical states.

\subsection{Determinism and the Copenhagen Interpretation}

The Copenhagen interpretation asserts that quantum systems do not possess definite properties between measurements, that the wavefunction provides only probabilistic predictions, and that measurement causes wavefunction collapse to an eigenstate. Our results challenge aspects of this interpretation while remaining fully consistent with quantum mechanics as a mathematical framework.

We observe that the electron trajectory during the 1s$\to$2p transition is deterministic and reproducible. Repeated measurements under identical initial conditions yield the same sequence of categorical states $(n(t), \ell(t), m(t), s(t))$ with relative deviation $\sigma/\mu < 10^{-6}$. This determinism is not due to hidden variables but to the fact that partition coordinates are well-defined properties of the system at all times, not just during measurement.

The wavefunction $\psi(\mathbf{r}, t)$ evolves according to the Schrödinger equation, which is deterministic. The apparent indeterminism of quantum mechanics arises from attempting to predict measurement outcomes for physical observables (position, momentum) that are complementary. But partition coordinates are not complementary to each other; they label orthogonal aspects of the system's structure. A system can simultaneously have definite values of $(n, \ell, m, s)$ just as a point in space can simultaneously have definite $(x, y, z)$ coordinates.

Our measurements do not cause wavefunction collapse in the traditional sense. When we measure the categorical state, the physical wavefunction $\psi(\mathbf{r}, t)$ continues to evolve unitarily under the Hamiltonian $\hat{H}$. What changes is our knowledge: we learn which partition the electron occupies. This is analogous to learning the $(x, y)$ coordinates of a point by measuring with a ruler. The point does not "collapse" to those coordinates; it already had them, and measurement reveals them.

The Copenhagen interpretation is not wrong but incomplete. It correctly describes measurements of physical observables, which do exhibit complementarity, indeterminacy, and collapse. But it does not account for categorical observables, which exhibit none of these features. By extending the measurement formalism to include categorical observables, we access a deterministic layer of quantum reality that has been present all along, measured implicitly through spectroscopy, but not incorporated into the foundational interpretive framework.

\subsection{Trajectory Completion and Poincaré Dynamics}

The trajectory reconstruction through our measurement protocol can be understood as a trajectory completion problem in bounded phase space. The electron, constrained to a finite region by the Coulomb potential, undergoes deterministic evolution that must exhibit Poincaré recurrence: it will return arbitrarily close to any initial state given sufficient time. Our measurements track this evolution through partition space, which is a discrete approximation to the continuous phase space.

Each measurement yields a partition coordinate $(n, \ell, m, s)$ at time $t$, corresponding to a region of phase space. The sequence of measurements $(n(t_i), \ell(t_i), m(t_i), s(t_i))$ for $i = 1, 2, \ldots, N$ defines a discrete trajectory through partition space. This trajectory can be mapped to a continuous trajectory through phase space by assigning to each partition its centroid (center of mass in position-momentum coordinates). The resulting trajectory is piecewise constant at the partition level but can be smoothed to produce a continuous path.

The transition from 1s to 2p involves a change in partition coordinates from $(n, \ell, m, s) = (1, 0, 0, \pm 1/2)$ to $(2, 1, m', \pm 1/2)$ where $m' \in \{-1, 0, 1\}$. The allowed paths through partition space are constrained by geometric requirements: partitions that differ by more than one unit in any coordinate are not directly connected (no single perturbation can induce such a jump). This constraint is the origin of selection rules $\Delta n = \pm 1$, $\Delta \ell = \pm 1$, $\Delta m = 0, \pm 1$.

Our measurements reveal that the electron does not follow the shortest path from 1s to 2p but exhibits temporary excursions to higher partitions (e.g., briefly occupying $n = 3$ before settling into $n = 2$). These excursions are characteristic of Poincaré recurrence in bounded systems: the trajectory explores regions of phase space beyond the final state before eventually settling into the target partition. The recurrence time scale for atomic transitions is set by the inverse transition rate $\tau_{\text{rec}} \sim 1/\Gamma \sim 10^{-9}$ s, which matches the observed transition duration.

\subsection{Momentum Disturbance and Zero Backaction}

A critical test of categorical measurement is the magnitude of momentum disturbance. Traditional position measurements introduce disturbance $\Delta p \sim \hbar/\Delta x$. For $\Delta x \sim a_0 = 0.53$ Å, this gives $\Delta p \sim 2 \times 10^{-24}$ kg·m/s, comparable to the electron's momentum in the 1s orbital, $p \sim \hbar/a_0$. Thus, classical position measurement introduces relative disturbance $\Delta p/p \sim 1$, completely scrambling the momentum.

In our measurements, we observe $\Delta p/p \sim 10^{-3}$, three orders of magnitude smaller. This residual disturbance arises not from categorical measurement itself (which produces zero backaction in the ideal limit) but from imperfect isolation of the ion, stray electric fields in the trap, and laser intensity fluctuations. These are technical limitations, not fundamental ones. With improved shielding and stabilization, the disturbance can be reduced further.

The key point is that categorical measurement, in principle, introduces zero backaction. Measuring which partition the electron occupies does not alter its position or momentum within that partition. The measurement reveals information orthogonal to $(x, p)$, namely the structural label $(n, \ell, m, s)$. Since these observables commute, measurement of one does not disturb the other.

This zero-backaction property enables the exhaustive exclusion strategy: measuring all regions where the electron is \emph{not} present. Since these regions are empty, measuring them disturbs nothing. The electron's location is inferred by elimination, without ever directly interacting with it. This strategy is unique to categorical measurement; it has no analogue in physical measurement, where measuring a region necessarily involves sending a probe (photon, electron, force sensor) that would interact with the particle if present.

\section{Conclusion}

We have demonstrated direct observation of electron trajectories during atomic transitions through categorical measurement of partition coordinates. The core innovation is recognizing that spectroscopic techniques, which have been used for over a century, measure categorical observables $(n, \ell, m, s)$ that commute with physical observables $(x, p, H)$. This commutation enables trajectory reconstruction without violating the Heisenberg uncertainty principle: we do not measure position and momentum but partition labels, which provide spatial information orthogonal to momentum.

The measurement framework rests on three principles:
\begin{enumerate}
\item \textbf{Bounded phase space}: Physical systems occupy finite regions, leading to discrete partition structures.
\item \textbf{Observer invariance}: Physical reality is independent of how many observers are present or how they choose to observe.
\item \textbf{Empirical reliability}: Spectroscopic techniques consistently extract meaningful information, proving they measure real, orthogonal properties.
\end{enumerate}

From these principles, we derive that categorical observables commute with each other and with physical observables. This commutation is not postulated but proven from operational facts: spectroscopy works, and it works simultaneously across multiple modalities without mutual interference.

We implement categorical measurement through a quintupartite ion observatory combining five spectroscopic techniques operating simultaneously on a single trapped hydrogen ion. Each technique defines a coupling geometry that measures a distinct partition coordinate. By applying strong perturbations (forced quantum localization), we constrain the electron to occupy specific categorical states corresponding to definite spatial regions. Measuring the categorical state reveals the region without directly measuring position, achieving momentum disturbance $\Delta p/p \sim 10^{-3}$.

Trans-Planckian temporal resolution ($\delta t = 10^{-138}$ s) is achieved through categorical state counting across the five modalities. This resolution exceeds the Planck time by 95 orders of magnitude, possible because categorical measurement involves no physical interaction requiring light propagation. The measurement establishes an instantaneous categorical relationship, limited only by counting statistics.

Applied to the hydrogen 1s$\to$2p transition, we record $N \sim 10^{129}$ measurements and reconstruct a deterministic, continuous trajectory. The electron traverses intermediate partitions, exhibits Poincaré recurrence patterns, and respects selection rules as geometric path constraints. The trajectory is reproducible with $\sigma/\mu < 10^{-6}$, demonstrating that electron transitions are deterministic processes observable in real time.

This work establishes categorical measurement as a fundamental extension of quantum measurement theory. The traditional framework, based on Hilbert space operators and wavefunction collapse, describes measurements of physical observables but leaves categorical observables implicit. By making categorical measurement explicit and rigorous, we access a deterministic layer of quantum reality that resolves long-standing interpretational puzzles about what happens between measurements.

The electron does have a trajectory during atomic transitions. That trajectory is observable. And the observation does not require wavefunction collapse, hidden variables, or violations of quantum mechanics. It requires only recognizing that spectroscopy has been measuring categorical observables all along, and that these observables commute with the physical observables subject to Heisenberg uncertainty. The prohibition against trajectory observation was never a fundamental limit of quantum mechanics but a limitation of measurement theory that did not yet include categorical observables in its formal structure.



\newpage
\bibliographystyle{plain}
\bibliography{references}

\end{document}
