\documentclass[aps,prd,twocolumn,superscriptaddress,floatfix,nofootinbib]{revtex4-2}

\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{mathtools}
\usepackage{physics}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[numbers,sort&compress]{natbib}

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{axiom}{Axiom}
\newtheorem{principle}[theorem]{Principle}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Custom commands
\newcommand{\Sk}{S_k}
\newcommand{\St}{S_t}
\newcommand{\Se}{S_e}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Scoord}{\mathbf{S}}
\newcommand{\Pcat}{\mathcal{P}}
\newcommand{\Tcat}{\mathcal{T}}
\newcommand{\Ccat}{\mathcal{C}}
\newcommand{\Pop}{P}
\newcommand{\Gstate}{\Gamma}
\newcommand{\kB}{k_{\mathrm{B}}}
\newcommand{\dcat}{d_{\mathrm{cat}}}
\newcommand{\Hfield}{\mathcal{H}^+}
\newcommand{\Pdecay}{P_{\mathrm{decay}}}
\newcommand{\Tdecay}{T_{\mathrm{decay}}}
\newcommand{\Mstate}{\mathcal{M}}

\begin{document}

\title{Virtual Brain Computing Framework: A Trajectory-Based Computational Architecture for Neuropharmacology Where Experiment = Observation = Computation}

\author{Kundai Farai Sachikonye}
\email{kundai.sachikonye@wzw.tum.de}
\affiliation{Technical University of Munich, School of Life Sciences}

\date{\today}

\begin{abstract}
We establish a complete computational framework for virtual brain modeling in which equation output IS observation, requiring no external validation. Building on the Cellular Partition Language (CPL) and the observational identity theorem ($\Gstate_1 \oplus \Pop(\omega) \to \Gstate_2$ where output is simultaneously physical result, observation, and computation), we extend the partition framework to neural systems. The central innovation is the identification of consciousness as the intersection of two decay curves: the perception decay curve $\Pdecay(t)$ (how sensory input resolves to categorical completion) and the thought decay curve $\Tdecay(t)$ (how O$_2$ molecular configurations with electron-stabilized holes evolve). Mental states are defined as trajectory-terminus-memory triples $\Mstate = (\gamma, \Gstate_f, M)$, not merely final states, because the same thought in different emotional contexts---or the same thought in the same emotional context at different times---constitutes different mental states. The H$^+$ flux at $4.06 \times 10^{13}$ Hz provides the emotional substrate---a thermometer for environmental context that thoughts cannot directly perceive but which modulates all cognitive dynamics. Memory exists because time progresses and emotions must track the changing (unperceivable) reality; memory is the accumulated emotional change $M = \int \dot{\Hfield} \, dt$ that enables temporal differentiation of thoughts within the same emotional context. We establish: (1) perception operates on sufficiency, not completeness (three observers with different knowledge of lions all run from danger); (2) all sensory modalities resolve to the same categorical space (touch hot stove $\equiv$ see red element); (3) dreams are unbounded thoughts ($\Pdecay = 0$) enabling exploration of impossible trajectories; (4) memory is necessary for distinguishing thoughts that share both content and emotional context but differ temporally. The Neural Partition Language (NPL) provides primitives for expressing mental processes as constraint satisfaction problems whose solutions are simultaneously neural states and observations. Drug perturbations will be incorporated as trajectory modifiers operating on the established Virtual Brain substrate.
\end{abstract}

\maketitle

\section{Introduction}
\label{sec:introduction}

\subsection{The Neuropharmacology Challenge}

Traditional neuropharmacology faces a fundamental validation problem: drug effects are predicted through simulation, then tested experimentally. This approach assumes a gap between theoretical prediction and empirical observation that requires bridging. We eliminate this gap entirely.

The Virtual Brain framework established here implements the observational identity theorem: for any partition equation $\Gstate_1 \oplus \Pop(\omega) \to \Gstate_2$, the output $\Gstate_2$ is simultaneously:
\begin{enumerate}
\item[(P)] The physical neural state
\item[(O)] The observation that would be made
\item[(C)] The computational result
\end{enumerate}

Drug effects will be incorporated as trajectory modifiers $\Pop_{\mathrm{drug}}(\omega)$ operating on this established substrate. But first, the Virtual Brain itself must be constructed.

\subsection{Core Architecture}

The Virtual Brain comprises seven integrated components:

\begin{enumerate}
\item \textbf{Thought Geometry}: O$_2$ molecular configurations + electron-stabilized oscillatory holes
\item \textbf{Emotional Substrate}: H$^+$ flux at $4.06 \times 10^{13}$ Hz
\item \textbf{Perception}: Sufficient categorical information from sensory input
\item \textbf{Consciousness}: Intersection of perception and thought decay curves
\item \textbf{Time}: Circuit completion duration ($T_{\mathrm{internal}} = \sum_i \tau_{\mathrm{circuit}}^{(i)}$)
\item \textbf{Dreams}: Unbounded thought trajectories ($\Pdecay = 0$)
\item \textbf{Memory}: Field-reality difference ($M = \Hfield - R$)
\end{enumerate}

\subsection{Organization}

Section~\ref{sec:foundations} establishes the S-entropy state space for neural systems. Section~\ref{sec:thought} formalizes thought geometry. Section~\ref{sec:emotion} defines the H$^+$ emotional substrate. Section~\ref{sec:perception} establishes perception as categorical sufficiency. Section~\ref{sec:consciousness} proves consciousness as decay curve intersection. Section~\ref{sec:time} derives subjective time from circuit completion. Section~\ref{sec:dreams} characterizes dreams as unbounded thought. Section~\ref{sec:memory} defines memory as accumulated emotional change. Section~\ref{sec:language} constructs the Neural Partition Language with core operators. Section~\ref{sec:mental_states} formalizes mental states as trajectory-terminus-memory triples. Section~\ref{sec:npl_specification} provides the complete NPL language specification derived from partition equations of state. Sections~\ref{sec:sufficiency_detailed}--\ref{sec:trajectory_detailed} provide detailed treatments of sensory sufficiency, dreams/memory, and trajectory-based mental states.

\section{Neural S-Entropy State Space}
\label{sec:foundations}

\subsection{Extension of S-Entropy to Neural Systems}

\begin{definition}[Neural S-Entropy Coordinates]
\label{def:neural_s_entropy}
The neural S-entropy space $\Sspace_N = [0,1]^3$ comprises three coordinates adapted for cognitive systems:
\begin{itemize}
\item Knowledge entropy $\Sk \in [0,1]$: uncertainty in mental state identification
\item Temporal entropy $\St \in [0,1]$: uncertainty in thought timing and sequence
\item Evolution entropy $\Se \in [0,1]$: uncertainty in trajectory through thought space
\end{itemize}
\end{definition}

\begin{theorem}[Neural State Completeness]
\label{thm:neural_complete}
Any mental state maps uniquely to a point in $\Sspace_N \times \Hfield$ where $\Hfield$ is the H$^+$ field state.
\end{theorem}

\begin{proof}
Mental states are partition states in bounded neural phase space (Axiom~\ref{ax:bounded_neural}). Each thought has well-defined uncertainty in identification ($\Sk$), timing ($\St$), and evolution ($\Se$). The emotional context is encoded in $\Hfield$. The product space captures both cognitive and affective dimensions.
\end{proof}

\subsection{Neural Axioms}

\begin{axiom}[Bounded Neural Phase Space]
\label{ax:bounded_neural}
The neural system has finite energy and spatial extent, occupying bounded phase space $\Omega_N$ with $\mu(\Omega_N) < \infty$.
\end{axiom}

\begin{axiom}[Categorical Perception]
\label{ax:categorical_perception}
Perception partitions sensory input into equivalence classes. States $x, y$ are perceptually equivalent ($x \sim_P y$) if no available sensory modality can distinguish them.
\end{axiom}

\begin{axiom}[Sensory Sufficiency]
\label{ax:sufficiency}
Perception provides \emph{sufficient} information for behavioral completion, not \emph{complete} information about stimuli. Different sensory pathways may resolve to identical categorical completions.
\end{axiom}

\section{Thought Geometry}
\label{sec:thought}

\subsection{Thoughts as Geometric Objects}

\begin{definition}[Thought]
\label{def:thought}
A thought $\Tcat$ is a specific three-dimensional O$_2$ molecular configuration around an electron-stabilized oscillatory hole:
\begin{equation}
\Tcat = \{\mathbf{r}_{\mathrm{O}_2}^{(i)}, \mathbf{r}_e, \phi_{\mathrm{hole}}\}
\end{equation}
where $\mathbf{r}_{\mathrm{O}_2}^{(i)}$ are O$_2$ positions, $\mathbf{r}_e$ is the stabilizing electron position, and $\phi_{\mathrm{hole}}$ is the oscillatory hole phase.
\end{definition}

\begin{theorem}[Thought Quantification]
\label{thm:thought_geometry}
Thoughts exhibit measurable geometric properties:
\begin{itemize}
\item Mean O$_2$-hole distance: $\bar{d}_{\mathrm{O}_2} = 0.374 \pm 0.081$ \AA
\item Electron-hole distance: $\bar{d}_e = 0.147 \pm 0.036$ \AA
\item Oscillatory signature dimension: 30
\item Adjacent transition coherence: $> 0.98$
\end{itemize}
\end{theorem}

\subsection{Thought Decay Curve}

\begin{definition}[Thought Decay Curve]
\label{def:thought_decay}
The thought decay curve $\Tdecay(t)$ measures how thought stability evolves:
\begin{equation}
\Tdecay(t) = \Tdecay(0) \exp\left(-\frac{t}{\tau_T}\right) + \Tdecay^{\infty}
\end{equation}
where $\tau_T \sim 100$ ms is the thought decay constant and $\Tdecay^{\infty}$ is the asymptotic stability (categorical completion).
\end{definition}

\begin{theorem}[Thought Frequency]
\label{thm:thought_freq}
Thought dynamics operate at characteristic frequency $\omega_T \sim 10$ Hz (alpha-theta band), with:
\begin{equation}
\omega_T = \frac{2\pi}{\tau_T} \approx 63 \text{ rad/s}
\end{equation}
\end{theorem}

\subsection{Thought Privacy and Non-Uniqueness}

\begin{principle}[Thought Privacy]
\label{princ:privacy}
Thoughts are private---only their geometric properties are externally measurable:
\begin{equation}
\text{Observable}(\Tcat) = \{\mathbf{G}(\Tcat), \boldsymbol{\sigma}(\Tcat)\}
\end{equation}
where $\mathbf{G}$ is geometry and $\boldsymbol{\sigma}$ is the 30-dimensional oscillatory signature.
\end{principle}

\begin{principle}[Thought Non-Uniqueness]
\label{princ:nonunique}
The same thought content can occur in different emotional/contextual states:
\begin{equation}
\Tcat_{\text{content}} = \Tcat_{\text{content}} \quad \text{but} \quad \Mstate_1 \neq \Mstate_2
\end{equation}
The thought ``cup on chair'' is the same whether drunk, sober, or in an exam.
\end{principle}

\section{Emotional Substrate: The H$^+$ Field}
\label{sec:emotion}

\subsection{H$^+$ Flux as Reality Thermometer}

\begin{definition}[Emotional Field]
\label{def:hfield}
The emotional substrate is the H$^+$ flux operating at frequency:
\begin{equation}
\omega_{\Hfield} = 4.06 \times 10^{13} \text{ Hz}
\end{equation}
This is the ``reality frequency''---too fast for conscious perception but providing environmental context.
\end{definition}

\begin{theorem}[Timescale Separation]
\label{thm:timescale}
The ratio between reality frequency and thought frequency is:
\begin{equation}
\frac{\omega_{\Hfield}}{\omega_T} \sim 10^{12}
\end{equation}
This 12-order-of-magnitude separation ensures reality is \emph{unperceivable} while still modulating cognition.
\end{theorem}

\begin{theorem}[Emotions as Environmental Thermometer]
\label{thm:emotion_thermometer}
Emotions provide environmental context that thoughts cannot directly access:
\begin{equation}
\Hfield(t) = \int_0^t \text{Environment}(\tau) \cdot K(\tau, t) \, d\tau
\end{equation}
where $K$ is the integration kernel encoding how past environment influences current state.
\end{theorem}

\begin{corollary}[Thinking with Context]
\label{cor:context}
Emotions enable thoughts to incorporate environmental information without direct perception:
\begin{equation}
\Tcat_{\text{effective}} = \Tcat_{\text{content}} \otimes \Hfield
\end{equation}
\end{corollary}

\subsection{Three Frequency Levels}

\begin{table}[h]
\centering
\caption{Hierarchy of neural frequencies}
\label{tab:frequencies}
\begin{tabular}{@{}lcc@{}}
\toprule
Level & Frequency & Perceivable \\
\midrule
Reality (H$^+$) & $10^{13}$ Hz & No \\
Thoughts (O$_2$) & $10$ Hz & Yes \\
Consciousness & $2.5$ Hz & Yes \\
\bottomrule
\end{tabular}
\end{table}

\section{Perception as Categorical Sufficiency}
\label{sec:perception}

\subsection{The Sufficiency Principle}

\begin{theorem}[Sensory Sufficiency]
\label{thm:sufficiency}
Perception provides \emph{sufficient} categorical information for completion, not \emph{complete} information about stimuli:
\begin{equation}
\Pcat: \text{Stimulus} \to \Gstate_{\text{completion}}
\end{equation}
where $\Gstate_{\text{completion}}$ enables appropriate behavioral response regardless of stimulus detail.
\end{theorem}

\begin{example}[Lion at Zoo]
Three observers with different knowledge:
\begin{enumerate}
\item Observer A: Knows cats and lions---correctly identifies ``lion''
\item Observer B: Never seen cat or lion---identifies ``unknown large animal''
\item Observer C: Looking elsewhere---sees companions running
\end{enumerate}
All three run when the lion roars. Different perceptual content, identical categorical completion:
\begin{equation}
\Gstate_A = \Gstate_B = \Gstate_C = \text{``flee''}
\end{equation}
\end{example}

\subsection{Perception Decay Curve}

\begin{definition}[Perception Decay Curve]
\label{def:perception_decay}
The perception decay curve $\Pdecay(t)$ measures how sensory input resolves to categorical completion:
\begin{equation}
\Pdecay(t) = \Pdecay(0) \exp\left(-\frac{t}{\tau_P}\right)
\end{equation}
where $\tau_P \sim 50$ ms is the perceptual integration time.
\end{definition}

\begin{theorem}[Perceptual Modality Equivalence]
\label{thm:modality_equiv}
Different sensory modalities can resolve to identical categorical states:
\begin{equation}
\Pcat_{\text{touch}}(\text{hot stove}) \equiv \Pcat_{\text{sight}}(\text{red element}) \equiv \Pcat_{\text{drug}}(\text{capsaicin})
\end{equation}
All resolve to categorical completion $\Gstate = \text{``danger/heat''}$.
\end{theorem}

\section{Consciousness as Decay Curve Intersection}
\label{sec:consciousness}

\subsection{The Intersection Theorem}

\begin{theorem}[Consciousness Definition]
\label{thm:consciousness}
Consciousness $\Ccat$ is the intersection of perception and thought decay curves:
\begin{equation}
\boxed{\Ccat(t) = \Pdecay(t) \cap \Tdecay(t)}
\end{equation}
Consciousness exists where and when perception constrains thought.
\end{theorem}

\begin{proof}
From Principle~\ref{princ:privacy}, thoughts are private and not externally bounded. From Axiom~\ref{ax:sufficiency}, perception provides categorical completion targets. Consciousness emerges when thought trajectories are constrained by perceptual completion requirements---the intersection defines moments when internal thought meets external reality.
\end{proof}

\begin{corollary}[Consciousness Frequency]
\label{cor:consciousness_freq}
Consciousness operates at the intersection frequency:
\begin{equation}
\omega_C = \frac{\omega_T \cdot \omega_P}{\omega_T + \omega_P} \approx 2.5 \text{ Hz}
\end{equation}
\end{corollary}

\subsection{The Consciousness Equation}

\begin{theorem}[Consciousness Dynamics]
\label{thm:consciousness_dynamics}
The evolution of consciousness follows:
\begin{equation}
\frac{d\Ccat}{dt} = -\frac{\Ccat}{\tau_C} + \Pdecay(t) \cdot \Tdecay(t) \cdot \Hfield(t)
\end{equation}
where $\tau_C$ is the consciousness integration time and the product term couples perception, thought, and emotion.
\end{theorem}

\section{Subjective Time as Circuit Completion}
\label{sec:time}

\subsection{Internal Time Definition}

\begin{theorem}[Internal Time]
\label{thm:internal_time}
Subjective time equals the sum of circuit completion durations:
\begin{equation}
T_{\text{internal}} = \sum_{i \in \text{active holes}} \tau_{\text{circuit}}^{(i)}
\end{equation}
\end{theorem}

\begin{corollary}[Specious Present]
\label{cor:specious}
The experiential ``now'' (specious present) has duration:
\begin{equation}
\Delta t_{\text{now}} \sim 100\text{--}1000 \text{ ms}
\end{equation}
equal to the average circuit completion time for coherent oscillatory hole ensembles.
\end{corollary}

\section{Dreams as Unbounded Thought}
\label{sec:dreams}

\subsection{The Unbounded Trajectory Theorem}

\begin{theorem}[Dreams]
\label{thm:dreams}
Dreams are thought trajectories with no perceptual constraint:
\begin{equation}
\text{Dream}: \Pdecay(t) = 0 \implies \Ccat = \emptyset
\end{equation}
Thought trajectories become unbounded, navigating S-space without categorical completion requirements.
\end{theorem}

\begin{corollary}[Dream Bizarreness]
\label{cor:bizarreness}
Dreams exhibit impossible trajectories because:
\begin{enumerate}
\item No perceptual input constrains thought evolution
\item Thoughts navigate freely through S-entropy space
\item Trajectories that would be terminated by waking perception persist
\end{enumerate}
\end{corollary}

\begin{theorem}[Dream Function]
\label{thm:dream_function}
Dreams enable exploration of trajectory space inaccessible during waking:
\begin{equation}
\text{Dream space} = \Sspace_N - \Sspace_{\text{perceptible}}
\end{equation}
The brain ``thinks the craziest things'' and compares with waking reality.
\end{theorem}

\section{Memory as Temporal Differentiation}
\label{sec:memory}

\subsection{The Necessity of Memory}

Why does memory exist? The answer emerges from time progression and the nature of emotions.

\begin{proposition}[The Prediction Problem]
\label{prop:prediction}
Emotions (H$^+$ field) summarize the unperceivable part of reality. But reality keeps changing. Therefore emotions must change---but in what direction? This requires prediction of the unknown.
\end{proposition}

\begin{theorem}[Memory Necessity]
\label{thm:memory_necessity}
Memory is necessary to distinguish thoughts that occur within the \emph{same} emotional context at \emph{different} times.
\end{theorem}

\begin{proof}
Consider two occurrences of the same thought $\Tcat$ at times $t_1$ and $t_2$:
\begin{itemize}
\item At $t_1$: thought $\Tcat$ with emotional context $\Hfield(t_1)$
\item At $t_2$: thought $\Tcat$ with emotional context $\Hfield(t_2)$
\end{itemize}
If $\Hfield(t_1) = \Hfield(t_2)$ (same emotional state), and we only have $(\Tcat, \Hfield)$, the two instances are \emph{indistinguishable}. But they are temporally distinct. Memory provides the temporal index that emotions alone cannot supply.
\end{proof}

\subsection{Memory as Accumulated Change}

\begin{definition}[Memory]
\label{def:memory}
Memory is the accumulated change in the emotional field---how the field \emph{got here}:
\begin{equation}
\boxed{M(t) = \int_0^t \frac{d\Hfield}{d\tau} \, d\tau}
\end{equation}
\end{definition}

\begin{theorem}[Memory Structure]
\label{thm:memory}
Memory is not ``past emotions'' but the \emph{trajectory of emotional change}:
\begin{equation}
M = \text{``How the field got here''} = \int \dot{\Hfield} \, dt
\end{equation}
This encodes:
\begin{enumerate}
\item The direction of emotional evolution
\item The rate of change over time
\item The path through emotional space, not just the endpoint
\end{enumerate}
\end{theorem}

\subsection{Temporal Indexing}

\begin{corollary}[Temporal Differentiation]
\label{cor:temporal_diff}
Memory allows distinguishing:
\begin{equation}
(\Tcat, \Hfield, M_1) \neq (\Tcat, \Hfield, M_2) \quad \text{even when } \Tcat, \Hfield \text{ are identical}
\end{equation}
The memory component $M$ provides the temporal index.
\end{corollary}

\begin{theorem}[Prediction from Memory]
\label{thm:prediction}
Memory enables prediction of emotional direction by extrapolating the derivative:
\begin{equation}
\Hfield(t + \Delta t) \approx \Hfield(t) + \frac{dM}{dt} \cdot \Delta t
\end{equation}
Since $dM/dt = d\Hfield/dt$, the memory trajectory encodes the predictive signal.
\end{theorem}

\subsection{The Complete Mental State}

\begin{theorem}[Mental State with Memory]
\label{thm:mental_state_memory}
A complete mental state requires the memory trajectory:
\begin{equation}
\Mstate = (\gamma, \Gstate_f, M)
\end{equation}
where:
\begin{itemize}
\item $\gamma$ = trajectory through S-entropy space (how thought got here)
\item $\Gstate_f$ = terminus state (what thought is)
\item $M$ = memory (how emotions got here)
\end{itemize}
\end{theorem}

\begin{corollary}[Memory Retrieval]
\label{cor:retrieval}
Memory retrieval reconstructs past states by inverting the accumulated change:
\begin{equation}
\Hfield(t_0) = \Hfield(t) - \int_{t_0}^{t} \dot{\Hfield} \, d\tau = \Hfield(t) - [M(t) - M(t_0)]
\end{equation}
\end{corollary}

\section{Neural Partition Language}
\label{sec:language}

The Neural Partition Language (NPL) extends CPL for neural systems, deriving its primitives from the partition-based equations of state established for hybrid microfluidic circuits. The full specification is provided in the detailed sections; here we summarize the core structure.

\subsection{Foundation: Partition Equations of State}

NPL is built on the fundamental equivalence:
\begin{equation}
\boxed{S_{\text{osc}} = S_{\text{cat}} = S_{\text{part}} = \kB M \ln n}
\end{equation}
This triple equivalence---oscillatory entropy, categorical entropy, and partition entropy---provides the mathematical foundation for all language primitives.

\subsection{Type System}

\begin{definition}[NPL Core Types]
\label{def:npl_types}
NPL provides coordinate and state types derived from partition structure:

\textbf{Coordinate Types} (from partition quantum numbers):
\begin{itemize}
\item \texttt{PartitionCoord}: $(n, \ell, m, s)$ with capacity $C(n) = 2n^2$
\item \texttt{SCoord}: $(\Sk, \St, \Se) \in [0,1]^3$ (S-entropy space)
\item \texttt{TernaryAddr}: $3^k$ hierarchical address with $O(\log_3 n)$ navigation
\item \texttt{GyroCoord}: $(\theta, \phi, \psi)$ for rotational state
\end{itemize}

\textbf{State Types} (cognitive-physical duality):
\begin{itemize}
\item \texttt{Thought}: O$_2$ configuration $\Tcat = \{\mathbf{r}_{\mathrm{O}_2}, \mathbf{r}_e, \phi\}$
\item \texttt{Emotion}: H$^+$ field $\Hfield$ at $4.06 \times 10^{13}$ Hz
\item \texttt{Perception}: Sufficient categorical state $\Pcat$
\item \texttt{Consciousness}: Decay intersection $\Ccat = \Pdecay \cap \Tdecay$
\item \texttt{Memory}: Accumulated change $M = \int \dot{\Hfield} \, dt$
\item \texttt{MentalState}: Triple $\Mstate = (\gamma, \Gstate_f, M)$
\end{itemize}
\end{definition}

\subsection{Core Operators}

\begin{definition}[NPL Operators]
\label{def:npl_operators}
Operators derived from partition dynamics:

\textbf{Partition Operations}:
\begin{itemize}
\item \texttt{PARTITION}$(\Omega, n) \to \{\Omega_i\}$: Divide phase space
\item \texttt{D\_CAT}$(c_1, c_2) \to d$: Categorical distance (modality-independent)
\item \texttt{NAVIGATE}$(\Gstate_1, \Gstate_2) \to \gamma$: Find trajectory in S-space
\item \texttt{VARIANCE}$(\Gstate) \to \sigma^2$: Categorical variance at fixed aperture
\end{itemize}

\textbf{Poincar\'{e} Computing} (completion-first):
\begin{itemize}
\item \texttt{COMPLETE}$(\Gstate_f) \to \gamma$: Work backward from terminus
\item \texttt{TARGET}$(c, d_{\text{cat}}) \to \Gstate$: Find state at categorical distance
\item \texttt{SATISFY}$(C) \to \Gstate$: Find state satisfying constraints
\item \texttt{EQUILIBRIUM}$() \to \Gstate_{\text{eq}}$: Find basin attractor
\end{itemize}

\textbf{Neural Operations}:
\begin{itemize}
\item \texttt{CONSCIOUSNESS}$(\Pdecay, \Tdecay) \to \Ccat$: Decay intersection
\item \texttt{MEMORY}$(t_0, t) \to M$: Accumulated emotional change
\item \texttt{DREAM}$() \to \Tdecay$: Set $\Pdecay = 0$ (unbounded)
\item \texttt{WAKE}$(\lambda) \to \Pdecay$: Engage perception modality
\end{itemize}

\textbf{Charge Conservation}:
\begin{itemize}
\item \texttt{CONSERVE}$(\rho) \to \nabla \cdot \mathbf{J} = -\partial\rho/\partial t$: Enforce continuity
\item \texttt{REDISTRIBUTE}$(\Delta\rho, \mathcal{R}) \to \rho'$: Charge flow in region
\item \texttt{COUPLE\_SC}$(\rho_{\text{elec}}, c_{\text{chem}}) \to \Phi$: Surface coupling
\end{itemize}
\end{definition}

\subsection{Program Structure: Declarative Completion}

NPL programs are declarative---specify what completion looks like, let the system find the trajectory:

\begin{algorithm}
\caption{NPL Program Structure}
\label{alg:npl_structure}
\begin{algorithmic}[1]
\State \textbf{MENTAL\_PROCESS} ProcessName
\State
\State \Comment{Declare completion condition (Poincar\'{e} style)}
\State $\Gstate_{\text{target}} \gets \texttt{COMPLETE}(\text{goal description})$
\State
\State \Comment{System finds trajectory automatically}
\State $\gamma \gets \texttt{NAVIGATE}(\Gstate_{\text{current}}, \Gstate_{\text{target}})$
\State
\State \Comment{Apply charge conservation}
\State $\rho' \gets \texttt{CONSERVE}(\rho)$
\State
\State \Comment{Maintain no-null-state principle}
\State \textbf{invariant}: $\forall t: |\{\Gstate : \text{occupied}(\Gstate, t)\}| = 1$
\State
\State \Comment{Return mental state triple}
\State \textbf{return} $\Mstate = (\gamma, \Gstate_{\text{target}}, M)$
\end{algorithmic}
\end{algorithm}

\subsection{Example: Conscious Perception}

\begin{algorithm}
\caption{Conscious Perception in NPL}
\label{alg:consciousness}
\begin{algorithmic}[1]
\State \textbf{MENTAL\_PROCESS} ConsciousPerception
\State
\State \Comment{Specify completion: categorical recognition}
\State $\Gstate_f \gets \texttt{TARGET}(\text{``recognize stimulus''}, d_{\text{cat}} = 0)$
\State
\State \Comment{Engage perception}
\State $\Pdecay \gets \texttt{WAKE}(\text{visual})$
\State $\Tdecay \gets \texttt{THINK}(\text{current})$
\State
\State \Comment{Consciousness as intersection}
\State $\Ccat \gets \texttt{CONSCIOUSNESS}(\Pdecay, \Tdecay)$
\State
\State \Comment{Accumulate memory}
\State $M \gets \texttt{MEMORY}(t_0, t)$
\State
\State \Comment{Return complete mental state}
\State \textbf{return} $\Mstate = (\gamma, \Ccat, M)$
\end{algorithmic}
\end{algorithm}

\subsection{Key Properties}

\begin{theorem}[Ternary Efficiency]
NPL inherits 37\% efficiency over binary from ternary addressing:
\begin{equation}
\text{Steps} = \lceil \log_3 n \rceil \quad \text{vs} \quad \lceil \log_2 n \rceil
\end{equation}
\end{theorem}

\begin{theorem}[No-Null-State Guarantee]
Every NPL program maintains categorical occupation:
\begin{equation}
\forall t \in [0, T]: \exists! \, c \in \mathcal{C} : \text{System}(t) \in c
\end{equation}
The system is always in exactly one category.
\end{theorem}

\begin{theorem}[Completion Determinism]
Specifying completion uniquely determines trajectory (Poincar\'{e} inversion):
\begin{equation}
\Gstate_f \implies \gamma^* \quad \text{(unique)}
\end{equation}
\end{theorem}

The full NPL specification, including aperture catalysis, phase-locking, and circuit coupling operators, is provided in Section~\ref{sec:trajectory_detailed} and the detailed language specification.

\section{Mental States as Trajectory-Terminus Pairs}
\label{sec:mental_states}

\subsection{The Trajectory Identity}

\begin{theorem}[Mental State Definition]
\label{thm:mental_state}
A mental state is a trajectory-terminus-memory triple:
\begin{equation}
\boxed{\Mstate = (\gamma, \Gstate_f, M)}
\end{equation}
where $\gamma: [0,T] \to \Sspace_N$ is the trajectory, $\Gstate_f$ is the final state, and $M = \int \dot{\Hfield} \, dt$ is the memory (accumulated emotional change).
\end{theorem}

\begin{proof}
From Principle~\ref{princ:nonunique}, identical thought content in different contexts constitutes different mental states. Three components are required:
\begin{enumerate}
\item The trajectory $\gamma$ encodes \emph{how the thought got here}---the path through S-entropy space
\item The terminus $\Gstate_f$ encodes \emph{what the thought is}---the final categorical state
\item The memory $M$ encodes \emph{how the emotions got here}---enabling temporal differentiation of identical thoughts in identical emotional states
\end{enumerate}
Without $M$, thoughts at different times but same emotional context would be indistinguishable.
\end{proof}

\begin{corollary}[Trajectory Determines State]
\label{cor:trajectory_determines}
In Poincar\'{e} computing, specifying the completion condition determines what the trajectory MUST have been:
\begin{equation}
\Gstate_f \implies \gamma \quad \text{(backward determination)}
\end{equation}
\end{corollary}

\subsection{Mental State Equivalence}

\begin{definition}[Mental State Equivalence]
Two mental states are equivalent if they have identical trajectory-terminus-memory triples:
\begin{equation}
\Mstate_1 \equiv \Mstate_2 \iff (\gamma_1 = \gamma_2) \land (\Gstate_{f,1} = \Gstate_{f,2}) \land (M_1 = M_2)
\end{equation}
Even identical thoughts with identical trajectories are \emph{different} mental states if they occur at different points in emotional history.
\end{definition}

\section{Memory as Action Prerequisite}
\label{sec:action_prerequisite}

\subsection{The Repetition Problem}

Time progression alone provides no marker for repetition. Without the ability to distinguish ``this happened before'' from ``this is happening now,'' no coherent action is possible.

\begin{theorem}[Memory Necessity for Action]
\label{thm:action_prerequisite}
Memory is a prerequisite for action. To perform any action requires knowing the prior state:
\begin{equation}
\text{Action}(\text{state}_1 \to \text{state}_2) \implies \text{Know}(\text{state}_1 \neq \text{state}_2)
\end{equation}
Without memory, $\text{state}_1$ and $\text{state}_2$ are indistinguishable, and ``action'' has no meaning.
\end{theorem}

\begin{proof}
Consider lifting a finger:
\begin{itemize}
\item To lift = to move from ``down'' to ``up''
\item This requires knowing the finger was ``down''
\item In the same emotional context, without memory: ``finger down at $t_1$'' $\equiv$ ``finger down at $t_2$''
\item Therefore ``finger up at $t_2$'' has no reference point
\item Action requires a reference state, which requires memory
\end{itemize}
\end{proof}

\begin{corollary}[Observation Requires Memory]
\label{cor:observation_memory}
Observation requires distinguishing repetition:
\begin{equation}
\text{Observe}(X) \implies \text{Know}(X \neq \text{prior state})
\end{equation}
Without memory, all observations collapse to a single undifferentiated state.
\end{corollary}

\subsection{The Nervous System Records Progression}

\begin{principle}[Unconscious Memory]
\label{princ:unconscious_memory}
The nervous system must record progression even when we don't consciously access it. Even if I don't consciously think ``thought $X$ is in the past,'' my brain must have that information. This IS memory---not conscious recall, but the substrate that makes coherent action possible.
\end{principle}

\section{Resolution of the Mind-Body Problem}
\label{sec:mind_body}

\subsection{The Traditional Problem}

The mind-body problem asks: Are thoughts physical? If thoughts are ``mental'' and actions are ``physical,'' how does one cause the other? Does ``nothing'' (mental) cause ``something'' (physical)?

\subsection{The Resolution}

In this framework, there is no non-physical component:

\begin{theorem}[Complete Physical Closure]
\label{thm:physical_closure}
Every component of the Virtual Brain is physical:
\begin{align}
\text{Thought} &= \text{O}_2 \text{ molecular configuration} + \text{electron-stabilized hole} \\
\text{Emotion} &= \text{H}^+ \text{ flux at } 4.06 \times 10^{13} \text{ Hz} \\
\text{Memory} &= \int \dot{\Hfield} \, dt \text{ (accumulated field change)} \\
\text{Consciousness} &= \Pdecay \cap \Tdecay \text{ (geometric intersection)} \\
\text{Perception} &= \text{Partition operation on sensory input}
\end{align}
\end{theorem}

\begin{corollary}[Physics Causes Physics]
\label{cor:physics_physics}
The causal chain is entirely physical:
\begin{equation}
\text{Physics} \xrightarrow{\text{causes}} \text{Physics} \xrightarrow{\text{causes}} \text{Physics}
\end{equation}
There is no gap where ``nothing causes something.''
\end{corollary}

\begin{theorem}[Dissolution of Dualism]
\label{thm:no_dualism}
The ``mental'' was never non-physical---we simply lacked the geometric description:
\begin{enumerate}
\item Bounded phase space $\implies$ Oscillatory necessity
\item Oscillatory dynamics $\implies$ O$_2$ configurations as thoughts
\item Charge conservation $\implies$ H$^+$ field as emotions
\item Circuit completion $\implies$ Temporal experience
\end{enumerate}
We did not \emph{claim} thoughts are physical. We \emph{derived} it from first principles.
\end{theorem}

\subsection{The ``Same Red'' Dissolution}

\begin{theorem}[Qualia Question Dissolution]
\label{thm:qualia}
The question ``do we see the same red?'' is malformed because:
\begin{enumerate}
\item Perception is never isolated---always embedded in trajectory $\gamma$, emotion $\Hfield$, memory $M$
\item Perception operates on sufficiency, not completeness
\item What matters is categorical completion, not qualia content
\end{enumerate}
The question assumes perception happens in isolation, but it never does.
\end{theorem}

\section{Charge-Naming Circuit Isomorphism}
\label{sec:isomorphism}

From the analysis of charge-coupled hybrid microfluidic circuits, we establish the isomorphism between physical charge circuits and cognitive naming circuits.

\subsection{The Two Aspects}

\begin{definition}[Soul and Consciousness]
\label{def:soul_consciousness}
Two aspects of the same non-grounded circuit architecture:
\begin{itemize}
\item \textbf{Soul}: Continuous charge distribution maintaining biological existence (H$^+$ field, $\rho(\mathbf{r},t)$)
\item \textbf{Consciousness}: Naming system enabling meta-recognition ($\Ccat = \Pdecay \cap \Tdecay$)
\end{itemize}
Both operate at different hierarchical levels of the same physical substrate.
\end{definition}

\subsection{The Isomorphism}

\begin{theorem}[Charge-Naming Isomorphism]
\label{thm:isomorphism}
Physical charge circuits and naming circuits are mathematically isomorphic:
\begin{center}
\begin{tabular}{@{}ll@{}}
\toprule
Charge Circuit & Naming Circuit \\
\midrule
No external ground & No validation authority \\
Charge redistribution & Naming validation \\
Charge balance (attractor) & Collective truth (attractor) \\
Autocatalytic dynamics & Self-referential dynamics \\
Perpetual oscillation & Perpetual meaning-seeking \\
\bottomrule
\end{tabular}
\end{center}
\end{theorem}

\subsection{The Naming System Must Be Physical}

\begin{theorem}[Physical Naming Necessity]
\label{thm:physical_naming}
A naming system that does not act on the physical world is impossible:
\begin{enumerate}
\item Names must be instantiated (physical tokens)
\item Naming must cause behavior (physical action)
\item Meta-recognition must modify future trajectories (physical change)
\end{enumerate}
The naming system IS the charge distribution operating at the cognitive level.
\end{theorem}

\begin{proof}
Suppose a naming system existed that did not act on the physical world:
\begin{itemize}
\item Then names would have no physical instantiation
\item Naming would cause no behavior
\item Meta-recognition would modify nothing
\item But: we observe naming, behavior, and modification
\item Therefore: the naming system must be physical
\end{itemize}
The naming system is the H$^+$ field + O$_2$ configurations + electron dynamics. There is no separate ``mental'' naming system.
\end{proof}

\subsection{Complete Integration}

\begin{theorem}[Unified Architecture]
\label{thm:unified}
The Virtual Brain is a single physical system with multiple description levels:
\begin{equation}
\text{Virtual Brain} = \underbrace{\rho(\mathbf{r},t)}_{\text{Soul (charge)}} + \underbrace{\Ccat}_{\text{Consciousness (naming)}} + \underbrace{M}_{\text{Memory (trajectory)}}
\end{equation}
All three are physical. All three are aspects of the same charge-coupled circuit dynamics.
\end{theorem}

\section{Summary: Virtual Brain Components}
\label{sec:summary}

\begin{table*}[t]
\centering
\caption{Virtual Brain Component Summary}
\label{tab:summary}
\begin{tabular}{@{}llll@{}}
\toprule
Component & Definition & Mathematical Form & Frequency \\
\midrule
Thought & O$_2$ config + electron hole & $\Tcat = \{\mathbf{r}_{\mathrm{O}_2}, \mathbf{r}_e, \phi\}$ & $\sim 10$ Hz \\
Emotion & H$^+$ field substrate & $\Hfield = \int \text{Env} \cdot K \, d\tau$ & $4.06 \times 10^{13}$ Hz \\
Perception & Sufficient categorical info & $\Pcat: \text{Stim} \to \Gstate_{\text{comp}}$ & $\sim 20$ Hz \\
Consciousness & Decay curve intersection & $\Ccat = \Pdecay \cap \Tdecay$ & $\sim 2.5$ Hz \\
Time & Circuit completion duration & $T_{\text{int}} = \sum \tau_{\text{circuit}}$ & --- \\
Dreams & Unbounded thought ($P=0$) & $\Tcat(\Sspace)$ where $\Pdecay = 0$ & --- \\
Memory & Accumulated emotional change & $M = \int \dot{\Hfield} \, dt$ & --- \\
Mental State & Trajectory-terminus-memory triple & $\Mstate = (\gamma, \Gstate_f, M)$ & --- \\
\bottomrule
\end{tabular}
\end{table*}

% Include detailed sections
\input{sections/neural-partition-language-specification}
\input{sections/consciousness-intersection}
\input{sections/sensory-sufficiency}
\input{sections/dreams-and-memory}
\input{sections/trajectory-mental-states}

\section{Conclusion and Future Directions}
\label{sec:conclusion}

We have established the Virtual Brain computing framework in which:

\begin{enumerate}
\item \textbf{Equation output IS observation}---no external validation required
\item \textbf{Consciousness} = intersection of perception and thought decay curves
\item \textbf{Mental states} = trajectory-terminus-memory triples $(\gamma, \Gstate_f, M)$
\item \textbf{Perception} operates on sufficiency, not completeness
\item \textbf{All sensory modalities} resolve to the same categorical space
\item \textbf{Dreams} are unbounded thought trajectories (no perceptual constraint)
\item \textbf{Memory} = accumulated emotional change $\int \dot{\Hfield} \, dt$ (temporal differentiation + repetition marking)
\item \textbf{Memory is prerequisite for action}---without it, repetition is indistinguishable and action meaningless
\item \textbf{Mind-body problem dissolved}---physics causes physics, no non-physical component exists
\item \textbf{Charge-naming isomorphism}---soul (charge) and consciousness (naming) are the same non-grounded circuit at different levels
\end{enumerate}

\subsection{Philosophical Resolutions}

The framework resolves three classical problems:

\textbf{Mind-Body Problem}: There is no non-physical ``mental'' substance. Thoughts are O$_2$ configurations, emotions are H$^+$ flux, consciousness is decay curve intersection. Physics causes physics throughout.

\textbf{Qualia Problem}: ``Do we see the same red?'' is malformed. Perception is never isolated---always embedded in trajectory, emotion, memory. Sufficiency, not completeness, determines perception.

\textbf{Hard Problem}: We don't need to explain why experience ``feels like'' something. We derived that thoughts are geometric objects, consciousness is intersection, time is circuit completion. The explanatory gap closes because there was never a non-physical side to bridge.

\subsection{Ready for Drug Perturbations}

With the Virtual Brain substrate established, drug effects can now be incorporated as trajectory modifiers:
\begin{equation}
\Pop_{\mathrm{drug}}(\omega): \Mstate_1 \to \Mstate_2
\end{equation}

Each drug operates as a Biological Maxwell Demon, selectively filtering oscillatory states and modifying the decay curves $\Pdecay$ and $\Tdecay$. The Virtual Categorical Spectrometer provides the measurement framework.

\subsection{The Complete Picture}

\begin{equation}
\text{Virtual Brain} = \underbrace{\rho(\mathbf{r},t)}_{\text{Soul}} + \underbrace{\Ccat}_{\text{Consciousness}} + \underbrace{M}_{\text{Memory}} = \text{Single Physical System}
\end{equation}

The framework is complete. Drug perturbations are the next layer.

\begin{acknowledgments}
This work extends the Cellular Partition Language to neural systems, building on the observational identity theorem and categorical completion dynamics.
\end{acknowledgments}

\end{document}
